---
title: "8b"
author: "Yiqiao Yin"
date: "2/23/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Simulation

Program a simulation in R related to the "Should we take measurements at an intermediate design point?" paper. Set up a scenario in which E(y|x) is a quadratic function of x, and where data could be taken at x=0,0.5,1 or just at 0 and 1, and compare these results for different levels of curvature. Unlike the published paper, use a binomial rather than a normal model, so that the outcome is the number of patients in each group who die.

Let us first define the data. We draw $n = 250$ samples from $\text{Bernoulli}(N, p = 0.5)$ distribution. We set the parameter $k = 1$ in *rbinom* function and this ensures that the samples only take values 1 or 0. In other words, we have the following:
$$X \sim \text{Bernoulli}(p = 0.5)$$
as the explanatory variables. As indicated in the problem, this ensures that we are modeling a binary variable.

Moreover, for some fixed values $a$, $b$, and $c$, we define a quadratic function as the following
$$y = aX^2 + bX + c$$
and since $X$ only takes values 1 or 0 we have outcome variable $y$ take only two values as well. 

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Create a N x k matrix of covariates
N = 250
X = cbind(rbinom(N, 1, 0.5)) # bernoulli: values take 1 or 0
print(table(X))

# Create a normally distributed variable that is a function of the covariates
a = 2
b = 3
c = 4
y = a*X^2 + b*X + c # y is defined as a quadratic function of x
print(table(y))
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* package
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan"
setwd(path)
file = path
set_cmdstan_path(file)
file <- file.path(path, "linearModel_8b.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Run *mod()* defined above
fit1 <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit1$summary()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
library(tidyverse)
fit1$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### II. Coefficients from Normal with Zero-center

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_ii.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Run *mod()* defined above
fit2 <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit2$summary()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Present traceplot for mixture and convergence
fit2$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### III. Coefficients from Double Exponential

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_iii.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Run *mod()* defined above
fit3 <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit3$summary()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
fit3$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### IV. Coefficients from Student-t

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_iv.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Run *mod()* defined above
fit4 <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit4$summary()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
fit4$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### LOO

Let us use leave-one-out to diagnose the fit for the above models.

```{r, messages=FALSE, warning=FALSE, error=FALSE}
loo2 <- fit2$loo()
loo3 <- fit3$loo()
loo4 <- fit4$loo()
```

```{r, messages=FALSE, warning=FALSE, error=FALSE}
print(loo2)
print(loo3)
print(loo4)
```