---
title: "4b"
author: "Yiqiao Yin"
date: "2/2/2021"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Problem 1:

Please see the previous part (written) part attached to this. 

I present the last part, i.e. the comparison of two estimates, within the restricted region of theta $[0,1]$ in the following.

```{r}
theta = seq(0, 1, 0.01)
plot(theta, theta^2 - theta + 1/3 - 1/12, col = as.numeric(theta^2 - theta + 1/3 - 1/12 < 0) + 2, 
     type = "p", cex = 0.5,
     xaxs = "i", yaxs = "i",
     xlab = "x: [0,1]", ylab = "difference",
     main = "theta1 - theta2")
```

### Problem 2: 

Consider a unit square. In other words, consider $X_1, X_2 \sim \text{uniform}([0,1])$ while $X_1$ and $X_2$ are independent. Then consider the following sequence with running index $n$:
$$
y := 
\left\{
\begin{array}{lcl}
4/n && \text{ if } (x_1^2 + x_2^2)^2 \le 1 \\
0 && \text{else } \\
\end{array}
\right.
$$

Let us simulate a unit square. Let us mark the dots $(X_1, X_2)$ that with distance less than or equal to 1 in one color and the rest of the dots in another color. We count the dots to be estimated area of a quarter of a circle. Then we recover estimated $\pi$ as $n \rightarrow \infty$.

```{r}
## Define a function to output a plot of pi
nRange <- seq(0, 1e3, 2e2)[-1]
pi_hat_vec <- rep(NA, length(nRange))
for (N in nRange) {
  x <- runif(N)
  y <- runif(N)
  d <- sqrt(x^2 + y^2)
  label <- ifelse(d < 1, 1, 0)
  pi_hat <- round(4*plyr::count(label)[2,2]/N,3)
  pi_hat_vec[which(N == nRange)] <- pi_hat
  par(mfrow=c(1,2))
  plot(
    x, y,
    col = label+1,
    main = paste0(
      "Simulation of Pi: N=", N,
      "; \nApprox. Value of Pi=", pi_hat),
    pch = 20, cex = 1)
  plot(
    nRange, pi_hat_vec, type = "both",
    main = paste0("Path for Simulated Pi; \nApprox. value = ", pi_hat)); 
  lines(nRange, y = rep(pi, length(nRange)))
}
```

Appendix:

Let me create the above simulation in more refined sequence of $n$'s. In the following, I will create a small animation, of which I will save as "GIF" format and upload as a separate file. 

```{r}
# # Library
# library(animation)
# 
# ## Plot Monte Carlo Simulation of Pi
# saveGIF({
#   ## Define a function to output a plot of pi
#   nRange <- seq(1e2, 1e4, 1e2)
#   pi_hat_vec <- rep(NA, length(nRange))
#   for (N in nRange) {
#     x <- runif(N)
#     y <- runif(N)
#     d <- sqrt(x^2 + y^2)
#     label <- ifelse(d < 1, 1, 0)
#     pi_hat <- round(4*plyr::count(label)[2,2]/N,3)
#     pi_hat_vec[which(N == nRange)] <- pi_hat
#     par(mfrow=c(1,2))
#     plot(
#       x, y,
#       col = label+1,
#       main = paste0(
#         "Simulation of Pi: N=", N,
#         "; \nApprox. Value of Pi=", pi_hat),
#       pch = 20, cex = 1)
#     plot(
#       nRange, pi_hat_vec, type = "both",
#       main = paste0("Path for Simulated Pi; \nApprox. value = ", pi_hat)); 
#     lines(nRange, y = rep(pi, length(nRange)))
#   }
# }, movie.name = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/4b/mc-sim-pi-adv.gif", interval = 0.8, nmax = 30, 
# ani.width = 480)
```