---
title: "6b"
author: "Yiqiao Yin"
date: "2/19/2021"
output:
  html_document:
    toc: yes
    toc_depth: 3
    toc_float: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Model Fitting and Expansion

Assignment:
Simulate data from a sparse regression where there are 100 predictors but most of the coefficients are near zero.  Then fit regression models in Stan with various hierarchical models for the coefficients, including 
- zero-centered normal ("ridge"), 
- double exponential ("lasso"), and 
- t models.

### I. Multivariate Regresson on Sparse Data

```{r, messages=FALSE, warning=FALSE}
# Create a N x k matrix of covariates
N = 250
K = 100

# Create covariates and column names
covariates = replicate(K, rbinom(n=N, 1, 0.5))
colnames(covariates) = paste0("X", 1:100)

# Create the model matrix with intercept
X = cbind(Intercept=1, covariates)

# Create the model matrix with intercept
X = cbind(Intercept=1, covariates)

# Create a normally distributed variable that is a function of the covariates
coefs = matrix(rnorm(100+1, 0, 1), 101)
mu = X %*% coefs
sigma = 2
y = rnorm(N, mu, sigma)
```

```{r, messages=FALSE, warning=FALSE}
# Run lm for later comparison; but go ahead and examine now if desired
modlm = lm(y~., data=data.frame(X[,-1]))
mean(modlm$coefficients)
```

```{r, messages=FALSE, warning=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE}
# Run *mod()* defined above
fit <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit$summary()
```

```{r, messages=FALSE, warning=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
library(tidyverse)
fit$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### II. Coefficients from Normal with Zero-center

```{r, messages=FALSE, warning=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_ii.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE}
# Run *mod()* defined above
fit <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit$summary()
```

```{r, messages=FALSE, warning=FALSE}
# Present traceplot for mixture and convergence
fit$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### III. Coefficients from Double Exponential

```{r, messages=FALSE, warning=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_iii.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE}
# Run *mod()* defined above
fit <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit$summary()
```

```{r, messages=FALSE, warning=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
fit$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```

### IV. Coefficients from Student-t

```{r, messages=FALSE, warning=FALSE}
# Create the data list object for Stan input
dat = list(N=N, K=ncol(X), y=y, X=X)

# Use Stan: here use *cmdstanr* pakcage
library(cmdstanr)
path = "C:/Users/eagle/OneDrive/Course/CU Stats/STATS GR6102 - Applied Statistics II/InClass/stan/"
file <- file.path(path, "linearModel_6b_iv.stan")
mod <- cmdstan_model(file)
mod$print()
```

```{r, messages=FALSE, warning=FALSE}
# Run *mod()* defined above
fit <- mod$sample(
  data = dat,
  seed = 123,
  chains = 4,
  parallel_chains = 4,
  save_warmup = TRUE,
  refresh = 0 )

# Report summary table
fit$summary()
```

```{r, messages=FALSE, warning=FALSE}
# Present traceplot for mixture and convergence
# ex: let us take the first 3 and the last 3 as example
library(bayesplot)
fit$draws()[,,c(1:3, 99:103)] %>% mcmc_trace()
```